---
title: "Inference for categorical data"
author: "Alexis Mekueko"
date: "10/9/2020"
output:
  html_document:
    df_print: paged
  beamer_presentation: default
  pdf_document: default
  slidy_presentation: default
header-includes:
- \usepackage{tikz}
- \usetikzlibrary{positioning,shapes.multipart,shapes}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Github Link: https://github.com/asmozo24/DATA606_Lab6





```{r load-packages, message=FALSE}

library(tidyverse) #loading all library needed for this assignment
library(openintro)
library(infer)
library(gplots)
#head(fastfood)
#library(readxl)
#library(data.table)
#library(readr)
#library(plyr)
#library(dplyr)
#library(dice)
# #library(VennDiagram)
# #library(help = "dice")
#library(DBI)
#library(dbplyr)

#library(rstudioapi)
#library(RJDBC)
#library(odbc)
#library(RSQLite)
#library(rvest)
#library(stringr)
#library(readtext)
#library(ggpubr)
#library(fitdistrplus)
#library(ggplot2)
#library(moments)
#library(qualityTools)
library(normalp)
#library(utils)
#library(MASS)
#library(qqplotr)
library(DATA606)
getLabs()
#library(StMoSim)



```



## The Data

You will be analyzing the same dataset as in the previous lab, where you delved into a sample from the Youth Risk Behavior Surveillance System (YRBSS) survey, which uses data from high schoolers to help discover health patterns. The dataset is called yrbss.


## Exercise 1

What are the counts within each category for the amount of days these students have texted while driving within the past 30 days?
Answer: There are about 8 categories if we remove the NA


```{r }

view(yrbss)

test_Wdrive<- yrbss %>% 
  count(text_while_driving_30d)
test_Wdrive


```


## Exercise 2


What is the proportion of people who have texted while driving every day in the past 30 days and never wear helmets?
Answer: I don't know if we should based this proportion on the total population or just the no_helmet1. Well, if we just want to look at the never wear helmet sample, then it make sense to say the proportion of is equal every day testing while driving and never wear helmet over the total never wear helmet population , p = 7.64%



```{r }

#let's cleanup the NA
is.na(yrbss) # checking if there is a missing data in the dataset, return is yes
sum(is.na(yrbss)) # file to big, checking the sum of all missing data 
yrbss1 <- na.omit(yrbss) # delete/remove the missings data because it is an imcomplete observation
yrbss1
view(yrbss1)

#Remember that you can use filter to limit the dataset to just non-helmet wearers. Here, we will name the dataset no_helmet.

no_helmet <- yrbss1 %>%
  filter(helmet_12m == "never")
no_helmet

# Also, it may be easier to calculate the proportion if you create a new variable that specifies whether the individual has texted every day while driving over the past 30 days or not. We will call this variable text_ind.

no_helmet1 <- no_helmet %>%
  mutate(text_ind = ifelse(text_while_driving_30d == "30", "yes", "no"))
no_helmet1
# want to see what the yes are: this filter could have been limited to "30" which represents the evey day in 30days, twd = text_while_driving_30d
no_helmet1_twd <- filter(no_helmet1, text_ind == "yes")
p <- nrow(no_helmet1_twd)/nrow(no_helmet1)
p

```


### Exercise 3

What is the margin of error for the estimate of the proportion of non-helmet wearers that have texted while driving each day for the past 30 days based on this survey?
Answer: In a confidence interval, z* x SE is called the margin of error, 0.8%, at 1000 repeat, lower_ci = 6.84% ; upper_ci = 8.54% 


```{r }

# When summarizing the YRBSS, the Centers for Disease Control and Prevention seeks insight into the population parameters. To do this, you can answer the question, “What proportion of people in your sample reported that they have texted while driving each day for the past 30 days?” with a statistic; while the question “What proportion of people on earth have texted while driving each day for the past 30 days?” is answered with an estimate of the parameter.
# 
# The inferential tools for estimating population proportion are analogous to those used for means in the last chapter: the confidence interval and the hypothesis test.

no_helmet1 %>%
  specify(response = text_ind, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

# 46% ± % 3% margin of error , 6.84 = 7.64 - margin_error = 0.8%
# assuming 95% confidence interval Z* = 1.96
n <- nrow(no_helmet1)
z <- 1.96
se <- sqrt((p*(1-p))/n)
margin_error <- z * se
margin_error




```
### Exercise 4

Using the infer package, calculate confidence intervals for two other categorical variables (you’ll need to decide which level to call “success”, and report the associated margins of error. Interpet the interval in context of the data. It may be helpful to create new data sets for each of the two countries first, and then use these data sets to construct the confidence intervals.
Answer: Let's assum teens of age <18  are very active , like every day in a week and get to sleep 8hrs. so I picked  physically_active_7d and school_night_hours_sleep...
What is the margin of error for the estimate of the proportion of physical active 7 days that get 8hrs of school night hours of sleep based on this survey? margin error is 1.8 

What is the margin of error for the estimate of the proportion of teens with physical active 7 days based on this survey ? 0.96% 
What is the margin of error for the estimate of the proportion of teens with 8hrs of school night hours of sleep based on this survey ? 0.88


```{r }

age<- yrbss1 %>% 
  count(age)

active_7d_total <- yrbss1 %>% 
  count(physically_active_7d )

active_7d <- yrbss1 %>%
  filter(physically_active_7d == "7")
active_7d

active_7d1 <- active_7d %>%
  mutate(text_ind2 = ifelse(school_night_hours_sleep == "8", "yes", "no")) 
active_7d1



active_7d <- yrbss1 %>%
  mutate(text_ind2 = ifelse(physically_active_7d == "7", "yes", "no"))
active_7d



active_7d1 %>%
  specify(response = text_ind2, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

active_7d_8hr <- filter(active_7d1, text_ind2 == "yes")

n <- nrow(active_7d1)
p <- nrow(active_7d_8hr)/nrow(active_7d)
z <- 1.96
se <- sqrt((p*(1-p))/n)
margin_error <- z * se
margin_error

#What is the margin of error for the estimate of the proportion of teans with physical active 7 days ?
active_7d <- yrbss1 %>%
  mutate(text_ind3 = ifelse(physically_active_7d == "7", "yes", "no"))

active_7d %>%
  specify(response = text_ind3, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

active_7d3 <- filter(active_7d, text_ind3 == "yes")

n <- nrow(active_7d)
p <- nrow(active_7d3)/nrow(active_7d)
p
z <- 1.96
se <- sqrt((p*(1-p))/n)
margin_error <- z * se
margin_error

# What is the margin of error for the estimate of the proportion of teens with 8hrs of school night hours of sleep based on this survey

teens8hr <- yrbss1 %>%
  mutate(text_ind4 = ifelse(school_night_hours_sleep == "8", "yes", "no"))

teens8hr %>%
  specify(response = text_ind4, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

teens8hr1 <- filter(teens8hr, text_ind4 == "yes")

n <- nrow(teens8hr)
p <- nrow(teens8hr1)/nrow(teens8hr)
p
z <- 1.96
se <- sqrt((p*(1-p))/n)
margin_error <- z * se
margin_error



```


### Exercise 5
 
Describe the relationship between p and me. Include the margin of error vs. population proportion plot you constructed in your answer. For a given sample size, for which value of p is margin of error maximized?
answer: the margin error increases as the sample proportion goes up to reach a peak at 50% , then margin of error drops to reach 0 as sample proportion reach out 100% of the population proportion.

```{r }

# How does the proportion affect the margin of error ?
# 
# Imagine you’ve set out to survey 1000 people on two questions: are you at least 6-feet tall? and are you left-handed? Since both of these sample proportions were calculated from the same sample size, they should have the same margin of error, right? Wrong! While the margin of error does change with sample size, it is also affected by the proportion.
# 
# Think back to the formula for the standard error: SE=sqrt(p(1−p)/n)
# . This is then used in the formula for the margin of error for a 95% confidence interval:
# ME=1.96×SE=1.96×sqrt(p(1−p)/n).
# Since the population proportion p is in this ME formula, it should make sense that the margin of error is in some way dependent on the population proportion. We can visualize this relationship by creating a plot of ME vs. p
# 
# .
# 
# Since sample size is irrelevant to this discussion, let’s just set it to some value (n=1000
# ) and use this value in the following calculations:

n <- 1000

# The first step is to make a variable p that is a sequence from 0 to 1 with each number incremented by 0.01. You can then create a variable of the margin of error (me) associated with each of these values of p using the familiar approximate formula (ME=2×SE).

p <- seq(from = 0, to = 1, by = 0.01)
me <- 2 * sqrt(p * (1 - p)/n)

# 
# Lastly, you can plot the two variables against each other to reveal their relationship. To do so, we need to first put these variables in a data frame that you can call in the ggplot function.

dd <- data.frame(p = p, me = me)
ggplot(data = dd, aes(x = p, y = me)) + 
  geom_line() +
  labs(x = "Population Proportion", y = "Margin of Error")



```

### Exercise 6
Describe the sampling distribution of sample proportions at n=300 and p=0.1. Be sure to note the center, spread, and shape.
Answer:  n = 300, p = 0.1 , np = 30 >10, n(1-p) = 270 > 10 so, normal distribution ,unimodal, centered at 0.1.

```{r }

# We have emphasized that you must always check conditions before making inference. For inference on proportions, the sample proportion can be assumed to be nearly normal if it is based upon a random sample of independent observations and if both np≥10 and n(1−p)≥10
# 
# . This rule of thumb is easy enough to follow, but it makes you wonder: what’s so special about the number 10?
# 
# The short answer is: nothing. You could argue that you would be fine with 9 or that you really should be using 11. What is the “best” value for such a rule of thumb is, at least to some degree, arbitrary. However, when np
# and n(1−p)
# 
# reaches 10 the sampling distribution is sufficiently normal to use confidence intervals and hypothesis tests that are based on that approximation.
# 
# You can investigate the interplay between n
# and p and the shape of the sampling distribution by using simulations. Play around with the following app to investigate how the shape, center, and spread of the distribution of p^ changes as n and p changes.



```

### Exercise 7
Keep n constant and change p. How does the shape, center, and spread of the sampling distribution vary as p changes. You might want to adjust min and max for the x-axis for a better view of the distribution.
Answer: keep n constant, the shape is normal and spread increases as p goes up to 50%, then spread decreases when p reaches 100%.



```{r }




```


### Exercise 8
Now also change n. How does n appear to affect the distribution of p^?
the distribution of p^ appears to be norma, unimodal with less spread as well as standard error decreases.


###More Practice

For some of the exercises below, you will conduct inference comparing two proportions. In such cases, you have a response variable that is categorical, and an explanatory variable that is also categorical, and you are comparing the proportions of success of the response variable across the levels of the explanatory variable. This means that when using infer, you need to include both variables within specify.



### Exercise 9

s there convincing evidence that those who sleep 10+ hours per day are more likely to strength train every day of the week? As always, write out the hypotheses for any tests you conduct and outline the status of the conditions for inference. If you find a significant difference, also quantify this difference with a confidence interval.
Tests with Two Independent Samples
H_null : teen sleeping 10+hours per day do not tell whether they train every day of the week.
H_alt : teen sleeping 10+hours per day do tell whether they train every day of the week
# decision rule, reject H_null if z_cal <-1.96 or z_call > 1.96...z_cal = 46.35 > 1.96
#I am a little sketicle since n1 = n2 , is it a pure coincidence or event sleeping 10+ and active every day are dependent?
# Conclusion: we reject the null hypothesis, so A 95% confidence interval for the corrolation in proportions of spleeing 10+hrs and active every day in a week is (25.49, 25.50)

```{r }


# We could compute this over the join event , like using the sleeping 10+ and active every day of the week
# we use the individual event

#What is the margin of error for the estimate of the proportion of teans with physical active 7 days ?
active_7d <- yrbss1 %>%
  mutate(text_ind3 = ifelse(physically_active_7d == "7", "yes", "no"))

active_7d %>%
  specify(response = text_ind3, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

active_7d3 <- filter(active_7d, text_ind3 == "yes")

n1 <- nrow(active_7d)
n1
x1 <- nrow(active_7d3)
x1
p1 <- nrow(active_7d3)/nrow(active_7d)
p1
z <- 1.96
se1 <- sqrt((p1*(1-p1))/n1)

#se <- sqrt((p*(1-p))/n)
# margin_error <- z * se
# margin_error

# What is the margin of error for the estimate of the proportion of teens with 8hrs of school night hours of sleep based on this survey

teens10hr <- yrbss1 %>%
  mutate(text_ind4 = ifelse(school_night_hours_sleep == "10+", "yes", "no"))

teens10hr %>%
  specify(response = text_ind4, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

teens10hr1 <- filter(teens10hr, text_ind4 == "yes")

n2 <- nrow(teens10hr)
n2
x2 <- nrow(teens10hr1)
x2
p2 <- nrow(teens10hr1)/nrow(teens10hr)
p2
z <- 1.96

p_all <- (x1+x2)/(n1+n2)
p_all
z_cal <- (p1-p2)/sqrt((p_all*(1-p_all))*((1/n1) +(1/n2)))
z_cal

se2 <- sqrt((p2*(1-p2))/n2)

margin_error
# CI p1-p2 _+ ZxSp(sqrt((1/n1)+(1/n2)), Sp= sqrt((((n1-1)*se1^2) + ((n2-1)*se2^2))/(n1+n2-2) )

Sp <- sqrt((((n1-1)*se1^2) + ((n2-1)*se2^2))/(n1+n2-2) )

margin_error<- z*Sp*(sqrt((1/n1)+(1/n2)))

low_ci <- p1-p2 - margin_error
low_ci
upper_ci <- p1-p2 + margin_error
upper_ci

# decision rule, reject H_null if z_cal <-1.96 or z_nul > 1.96
#I am a little sketicle since n1 = n2 , is it a pure coincidence or event sleeping 10+ and active every day are dependent?




```



### Exercise 10

Let’s say there has been no difference in likeliness to strength train every day of the week for those who sleep 10+ hours. What is the probablity that you could detect a change (at a significance level of 0.05) simply by chance? Hint: Review the definition of the Type 1 error.
answer: 
A Type 1 Error is rejecting the null hypothesis when H_null is actually true.
we do not want to incorrectly reject H_null more than 5% of the time. This corresponds to a significance level of
0.05. That is, if the null hypothesis is true, the significance level indicates how often the data lead us to incorrectly reject H_null, meaning there is 5% margin we could be wrong. We could detect the change by checking success-failure and computing SE for the hypothesis test. in this case , z_cal <-1.96 or z_cal > 1.96



### Exercise 11

Suppose you’re hired by the local government to estimate the proportion of residents that attend a religious service on a weekly basis. According to the guidelines, the estimate must have a margin of error no greater than 1% with 95% confidence. You have no idea what to expect for p. How many people would you have to sample to ensure that you are within the guidelines?
Hint: Refer to your plot of the relationship between p and margin of error. This question does not require using a dataset
Answer: I would have to sample 9604 people to ensure that you are within the guidelines

```{r }

# I think p was 50% ...it could be hard to set p
p <- 0.50
margin_error <- 0.01 
z <- 1.96
se <- sqrt((p*(1-p))/n)
# margin_error <- z * se
n <- ((z^2)*(p*(1-p))) / (margin_error^2)
n


```

